Technique for \earlywarning\ detection of compact binary coalescence and its implications for multi-messenger astronomy

### 2011-06-08 introduction.tex
next round of comments
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@975 ccdd05d5-56c1-461a-bf98-85f6876523b4

To this end, we have the ambition of reporting \GW{} candidates
not minutes \emph{after} the merger, as has already been accomplished, but
seconds \emph{before}.  By examining the signal to noise ratio (\SNR{}) stream
for threshold crossings before the \GW{} signal leaves the
detection band, it is possible to trade some \SNR{} and sky localization
accuracy~\cite{Fairhurst2009} for latency.  Figure \ref{fig:earlywarning} shows
projected early trigger rates for \textsc{ns}--\textsc{ns} binaries in Advanced
\LIGO\ assuming the event rate predictions in~\cite{Abadie:2010p10836}.
%
\begin{figure}
\begin{center}
\includegraphics{figures/snr_in_time.pdf}
\caption{\label{fig:earlywarning} Expected number of \textsc{ns}--\textsc{ns} sources that will
be detectable $t$ seconds before coalescence.  The heavy solid line is the most
likely yearly rate estimate $\dot N_{\mathrm{re}}$ during Advanced \LIGO.  The
light solid lines represent the confidence interval $[\dot N_{\mathrm{low}}, \dot 
N_{\mathrm{high}}]$ described in
\cite{Abadie:2010p10836}.  Advanced \LIGO\ noise projections are described in 
\cite{ALIGONoise}.  Assuming that an \SNR\ of 8 is sufficient for
detection and that we observe $\dot N_{\mathrm{re}} = 40$ events~yr$^{-1}$ with a
detector in the `zero detuning, high power' configuration, about 10~yr$^{-1}$ will be 
detectable within 10 seconds before merger and $\sim1$~yr$^{-1}$ will be detectable 
within 100 seconds of merger.  The `zero detuning, lower power' and `\textsc{bbh} optimized' configurations are still more conducive to advance detection
but at the cost of fewer total detectable events above an \SNR\ of 8.}
\end{center}
\end{figure}

>>>>

To this end, we have the ambition of reporting \GW{} candidates not minutes
\emph{after} the merger, as has already been accomplished, but seconds
\emph{before}.  By examining the signal to noise ratio (\SNR{}) stream for
threshold crossings before the \GW{} signal leaves the detection band, it is
possible to trade some \SNR{} and sky localization
accuracy~\cite{Fairhurst2009} for latency.  Figure \ref{fig:earlywarning} shows
projected early trigger rates for \textsc{ns}--\textsc{ns} binaries in Advanced
\LIGO\ assuming the event rate predictions in~\cite{Abadie:2010p10836}.  The
most likely estimates indicate that 10 sources could be audible (meaning
producing an \SNR\ above 8) 10 seconds or more prior to the merger during one
year of observation.  This number is highly uncertain.  In fact as many as 100
events detected 10 seconds early is plausible. It is also plausible that no
sources will be detected at all.  The gray bands give the pessimistic to
optimistic interval defined in ~\cite{Abadie:2010p10836}. 
%
\begin{figure}
\begin{center}
\includegraphics{figures/snr_in_time.pdf}
\caption{\label{fig:earlywarning} Expected number of \textsc{ns}--\textsc{ns}
sources that will be detectable $t$ seconds before coalescence.  The heavy
solid line is the most likely yearly rate estimate $\dot N_{\mathrm{re}}$
during Advanced \LIGO.  The shaded region represents the confidence interval
$[\dot N_{\mathrm{low}}, \dot N_{\mathrm{high}}]$ described in
\cite{Abadie:2010p10836}.  Advanced \LIGO\ noise projections are described in
\cite{ALIGONoise}.  Assuming that an \SNR\ of 8 is sufficient for detection and
that we observe $\dot N_{\mathrm{re}} = 40$ events~yr$^{-1}$ with a detector in
the `zero detuning, high power' configuration, about 10~yr$^{-1}$ will be
detectable within 10 seconds before merger and $\sim1$~yr$^{-1}$ will be
detectable within 100 seconds of merger.} 
\end{center}
\end{figure}
